Concurrent Programming with CXL

### 2020-06-23 book/paper.tex
Latest

      processes with private variables and exchanging messages}

@@

\section*{Exercises}
\begin{problems}
\item See what happens if, instead of initializing \texttt{result} to 0,
you initialize it to 1.  (You do not need to understand the error report at this time.)
Explain in English what the problem is.
\item Write a CXL program that computes squares by repeated adding.  So the program
should compute the square of $x$ by adding $x$ to an initial value of 0 $x$ times.
\end{problems}

@@

\section*{Exercices}
\begin{problems}
\item Add an assertion that demonstrates that at most three
Italian philosophers are eating at the same time?
\item Implement a test program that checks to see that
the solution allows more than one Italian philosopher to eat at the same time?
\item Figure~\ref{fig:qsort} shows an iterative implementation of the Qsort
algorithm, and Figure~\ref{fig:qsorttest} an accompanying test program.
The array to be sorted is stored in  shared variable \texttt{a}.
Another shared variable, \texttt{todo}, contains the ranges of the
array that need to be sorted (initially the entire array).
Re-using as much of this code as you can, implement a parallel version of
this.  You should not have to change the methods \texttt{swap}, \texttt{partition},
or \texttt{sortrange} for this.  Create \texttt{NWORKERS} ``worker processes''
that should replace the \texttt{qsort} code.  Each worker loops until \texttt{todo}
is empty and sorts the ranges that it finds until then.  The \texttt{main}
process needs to wait until all workers are done.  Use a semaphore for this:
the \texttt{main} process should use \texttt{P} to wait for each
worker to terminate.  After that it should check that the array is sorted and
that it is a permutation of the original array.
\end{problems}

@@

The producer/consumer pattern is common.  Processes may be arranged
in pipelines, where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers executing them in parallel.

@@

\chapter{Actors and Synchronized Queues}
\label{ch:actor}
\index{actor model}
\glsadd{actor model}

@@

Some programming languages favor a completely different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through message queues.
See Figure~\ref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.

@@

The \texttt{synch} library supports a synchronized message queue,
similar to the \texttt{Queue} object in Python.
Its interface is as follows:
\begin{itemize}
\item \texttt{Queue()} returns a new message queue;
\item \texttt{enqueue(q, item)} add \texttt{item} to the queue pointed to by \texttt{q};
\item \texttt{dequeue(q)} waits for and returns an item on the queue pointed to by \texttt{q}.
\end{itemize}

@@

Figure~\ref{fig:actor} shows a CXL solution to the producer/consumer problem using
an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \texttt{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \texttt{balance} and \texttt{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

>>>>

      processes with private variables that communicate through message passing}

@@

\section*{Exercises}
\begin{problems}
\item See what happens if, instead of initializing \texttt{result} to 0,
you initialize it to 1.  (You do not need to understand the error report at this time.  They will be explained in more detail in Chapter~\ref{ch:cxlmachine}.)
\item Write a CXL program that computes squares by repeated adding.  So the program
should compute the square of $x$ by adding $x$ to an initial value of 0 $x$ times.
\end{problems}

@@

\section*{Exercices}
\begin{problems}
\item Add an assertion that demonstrates that at most three
Italian philosophers are eating at the same time?
\item Implement a test program that checks to see that
the solution allows more than one Italian philosopher to eat at the same time?
\item \label{ex:qsort}
Figure~\ref{fig:qsort} shows an iterative implementation of the Qsort
algorithm, and Figure~\ref{fig:qsorttest} an accompanying test program.
The array to be sorted is stored in  shared variable \texttt{a}.
Another shared variable, \texttt{todo}, contains the ranges of the
array that need to be sorted (initially the entire array).
Re-using as much of this code as you can, implement a parallel version of
this.  You should not have to change the methods \texttt{swap}, \texttt{partition},
or \texttt{sortrange} for this.  Create \texttt{NWORKERS} ``worker processes''
that should replace the \texttt{qsort} code.  Each worker loops until \texttt{todo}
is empty and sorts the ranges that it finds until then.  The \texttt{main}
process needs to wait until all workers are done.  Use a semaphore for this:
the \texttt{main} process should use \texttt{P} to wait for each
worker to terminate.  After that it should check that the array is sorted and
that it is a permutation of the original array.
\end{problems}

@@

The producer/consumer pattern is common.  Processes may be arranged
in \emph{pipelines},
\index{pipeline}
where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers executing them in parallel.

@@

\chapter{Actors and Message Passing}
\label{ch:actor}
\index{actor model}
\index{message passing}
\glsadd{actor model}

@@

Some programming languages favor a different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through message passing.
See Figure~\ref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.

@@

The \texttt{synch} library supports a synchronized message queue,
similar to the \texttt{Queue} object in Python.
Its interface is as follows:
\begin{itemize}
\item \texttt{Queue()} returns a new message queue;
\item \texttt{enqueue(q, item)} adds \texttt{item} to the queue pointed to by \texttt{q};
\item \texttt{dequeue(q)} waits for and returns an item on the queue pointed to by \texttt{q}.
\end{itemize}

@@

Figure~\ref{fig:actor} shows a CXL solution to the producer/consumer problem
(with an unbounded buffer)
using an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \texttt{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \texttt{balance} and \texttt{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

\section*{Exercises}
\begin{problems}
\item Implement a version of parallel qsort (see Example~\ref{ex:qsort})
using the actor model.  Use an actor for each worker as well as a main
actor that waits for each worker actor to have finished.
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\end{problems}
